Global patterns of forest autotrophic carbon fluxes

### 2020-02-15 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd

**Carbon cycling in forests worldwide is central to sustaining an estimated #% of Earth's biodiversity (*REF*), while strongly influencing atmospheric CO~2~ and climate** ([Bonan 2008](https://science.sciencemag.org/content/320/5882/1444)). Forests' autotrophic carbon flux (FACF)--that is, carbon fixation, allocation, and metabolism by trees and other primary producers--sets the energy ultimately available to heterotrophic organisms (including microbes), in turn influencing their abundance (**REFS**) and possibly diversity [@waide_relationship_1999] (**REFS**). FACF influences all organic matter stocks in forest ecosystems and is linked to cycling of energy, water, and nutrients (**REFS**) [@piao_forest_2010]. Critically, FACF also defines forest interactions with Earth's changing climate. Approximately **##** Gt of CO~2~ cycle through Earth's forests each year **REF-KAT**), and in recent decades their net C sequestration (~2.4 GtCyr^-1^) offset roughly 30% of anthropogenic fossil fuel emissions [@pan_large_2011]. As atmospheric carbon dioxide levels continue to rise, driving climate change, forests will play a critical role in shaping the future of Earth's climate (** REFS**). However, our ability to draw general macroscopic conclusions regarding global variation in multiple FACF variables with respect to climate has been limited in that these analyses often mix forests that vary in stand age, disturbance history, and/or management status; do not always sufficiently parse related variables; and typically consider only one or a few variables at a time.

@@

\renewcommand{\arraystretch}{2}
```{r eval = TRUE, echo=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)
hypothesis_table <- read.csv("hypothesis_table.csv", stringsAsFactors = FALSE, check.names = FALSE)
kable(hypothesis_table, booktabs = TRUE, caption = "Summary of hypotheses, corresponding specific predictions, and results. Direction of significant relationships is indicated by '+' and '-', and  shape is summarized as linear (L), concave up (CU), and concave down (CD). 'n.s.' indicates that the relationship is not signficant.") %>%
  kable_styling(latex_options = c("scale_down", "hold_position"), font_size = 12) %>%
  pack_rows("Q1. How do FACFs vary with latitude?", 1, 1, latex_gap_space = "1em", colnum = 1, hline_before = FALSE) %>%
  pack_rows("Q2. How do FACFs vary with MAT and MAP?", 2, 4, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
  pack_rows("Q3. How are FACFs related to climate variables, other than MAT and MAP?", 5, 7, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
  pack_rows("Q4. How does seasonality influence FACFs?", 8, 11, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
  pack_rows("Q5. When standardised by growing season length, how do FACFs vary with climate?", 12, 15, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
  add_header_above(c(" ", " ", " ", "Forest autotrophic carbon fluxes (FACF)" = 9, " ")) %>%
  #column_spec(1, width = "15cm") %>%
  column_spec(2, width = "5cm") %>%
  column_spec(3:13, width = "1.8cm") %>%
  kableExtra::landscape()
```

@@

\newpage  
| Variable | Definition | Components included | Methodologies used | Number of records | Number of geographic areas |
|--------------|-------------------|-------------|------------------------|-------|-------|
| $GPP$ | Gross Primary Production | NA | flux partitioning of eddy-covariance; $NPP$+$R_{auto}$ | 243 | 49 |
| $NPP$ | Net Primary Production | stem, foliage, coarse root, fine root, optionally others (e.g., branch, reproductive, understory) | $ANPP$ + $BNPP$ (majority); $GPP$-$R_{auto}$ | 161 | 56 |
| $ANPP$ | Aboveground Net Primary Production | stem, foliage, optionally others (e.g., branch, reproductive, understory) |$ANPP_{woody-stem}$ + $ANPP_{foliage}$ (+ others) | 278 | 86 |
| $ANPP_{woody-stem}$ | Woody stem growth component of $ANPP$ | woody stems down to DBH $\le$ 10cm (no branch turnover) | stem growth measurements scaled to biomass using allometries | 264 | 96 |
| $ANPP_{foliage}$ | Foliage component of $ANPP$ | foliage | litterfall collection (separated into components) | 98 | 49 |
| $BNPP$  | Belowground Net Primary Production | coarse and fine roots |  coarse roots estimated indirectly using allometries based on aboveground stem increment measures ; fine roots as below | 101 | 48 |
| $BNPP_{fine.root}$ | Fine root component of $BNPP$ | fine roots | measurements combined one or more of the following: soil cores, minirhizotrons, turnover estimates,  root ingrowth cores | 88 | 41 |
| $R_{root}$ | Root Respiration | coarse and fine roots | partitioning of total soil respiration (e.g., through root exclusion), scaling of root gas exchange | 64 | 26 |
| $R_{auto}$ | Autotrophic Respiration | foliage, stem, and root | chamber measurements of component gas exchange | 22 | 13 |

>>>>

**Carbon cycling in forests worldwide is central to sustaining an estimated #% of Earth's biodiversity (*REF*), while strongly influencing atmospheric CO~2~ and climate** [@bonan_forests_2008]. Forests' autotrophic carbon flux (FACF)--that is, carbon fixation, allocation, and metabolism by trees and other primary producers--sets the energy ultimately available to heterotrophic organisms (including microbes), in turn influencing their abundance (**REFS**) and possibly diversity [@waide_relationship_1999] (**REFS**). FACF influences all organic matter stocks in forest ecosystems and is linked to cycling of energy, water, and nutrients (**REFS**) [@piao_forest_2010]. Critically, FACF also defines forest interactions with Earth's changing climate. Approximately **##** Gt of CO~2~ cycle through Earth's forests each year **REF-KAT**), and in recent decades their net C sequestration (~2.4 GtCyr^-1^) offset roughly 30% of anthropogenic fossil fuel emissions [@pan_large_2011]. As atmospheric carbon dioxide levels continue to rise, driving climate change, forests will play a critical role in shaping the future of Earth's climate (** REFS**). However, our ability to draw general macroscopic conclusions regarding global variation in multiple FACF variables with respect to climate has been limited in that these analyses often mix forests that vary in stand age, disturbance history, and/or management status; do not always sufficiently parse related variables; and typically consider only one or a few variables at a time.

@@

\renewcommand{\arraystretch}{2}
```{r eval = TRUE, echo=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)
hypothesis_table <- read.csv("hypothesis_table.csv", stringsAsFactors = FALSE, check.names = FALSE)
kable(hypothesis_table, booktabs = TRUE, caption = "Summary of research questions, corresponding hypotheses, and results.'yes' indicates the hypothesis is supported, 'no' indicates the hypothesis is not supported, and 'n.s.' indicates that the relationship is not signficant.") %>%
  kable_styling(latex_options = c("scale_down", "hold_position"), font_size = 12) %>%
  pack_rows("Q1. How do FACFs vary with latitude?", 1, 1, latex_gap_space = "1em", colnum = 1, hline_before = FALSE) %>%
  pack_rows("Q2. How do FACFs vary with MAT and MAP?", 2, 4, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
  pack_rows("Q3. How are FACFs related to other climate variables?", 5, 7, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
  pack_rows("Q4. How does seasonality influence FACFs?", 8, 11, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
  pack_rows("Q5. When standardised by growing season length, how do FACFs vary with climate?", 12, 15, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
  add_header_above(c(" ", " ", " ", "Forest autotrophic carbon fluxes (FACF)" = 9, " ")) %>%
  #column_spec(1, width = "15cm") %>%
  column_spec(2, width = "5cm") %>%
  column_spec(3:13, width = "1.8cm") %>%
  kableExtra::landscape()
```

@@

\newpage  
| Variable | Definition | Components included | Methodologies used | Number of records | Number of geographic areas |
|--------------|-------------------|-------------|------------------------|-------|-------|
| $GPP$ | Gross Primary Production | NA | flux partitioning of eddy-covariance; $NPP$+$R_{auto}$ | 243 | 49 |
| $NPP$ | Net Primary Production | stem, foliage, coarse root, fine root, optionally others (e.g., branch, reproductive, understory) | $ANPP$ + $BNPP$ (majority); $GPP$-$R_{auto}$ | 161 | 56 |
| $ANPP$ | Aboveground Net Primary Production | stem, foliage, optionally others (e.g., branch, reproductive, understory) |$ANPP_{woody-stem}$ + $ANPP_{foliage}$ (+ others) | 278 | 86 |
| $ANPP_{woody stem}$ | Woody stem growth component of $ANPP$ | woody stems down to DBH $\le$ 10cm (no branch turnover) | stem growth measurements scaled to biomass using allometries | 264 | 96 |
| $ANPP_{foliage}$ | Foliage component of $ANPP$ | foliage | litterfall collection (separated into components) | 98 | 49 |
| $BNPP$  | Belowground Net Primary Production | coarse and fine roots |  coarse roots estimated indirectly using allometries based on aboveground stem increment measures ; fine roots as below | 101 | 48 |
| $BNPP_{fine root}$ | Fine root component of $BNPP$ | fine roots | measurements combined one or more of the following: soil cores, minirhizotrons, turnover estimates,  root ingrowth cores | 88 | 41 |
| $R_{root}$ | Root Respiration | coarse and fine roots | partitioning of total soil respiration (e.g., through root exclusion), scaling of root gas exchange | 64 | 26 |
| $R_{auto}$ | Autotrophic Respiration | foliage, stem, and root | chamber measurements of component gas exchange | 22 | 13 |
